Generalized Tractability

### 2023-04-18 Older_versions/GeneralizedTractability_230411.tex
Fred M2 2023 Mac Updates

Let $h:[0,\infty) \to [1,\infty)]$ satisfy
\begin{equation} \label{eq:h_cond}
	h \text{ is strictly increasing}, \qquad 
	h(0)=1, \qquad  [h(x)]^\tau \le h(\tau x) \quad \forall x \ge 0, \tau > 1.
\end{equation}

\begin{definition} \label{def:subhT}
	A problem is sub-$h \circ T$ tractable if
	\begin{equation*}
		\lim_{\varepsilon^{-1} + d \to \infty} \frac{\comp(\varepsilon,d)}{h(cT(\varepsilon^{-1},d,\vp))} = 0 \qquad \text{\sout{$\forall c>0$}}
	\end{equation*}
\end{definition}

@@

\noindent \textbf{Sufficient Condition:}\\	
Suppose that $S_c <\infty$ for all $c>0$.  Since the terms of the series in the definition of $S_c$ are non-increasing, we have 
	\[
		\frac{n}{h(cT(\lambda_{n,d}^{-1},d))}\le S_c \qquad \forall n,d \in \naturals.
	\]
or equivalently
\[
	n\le S_c \, h(cT(\lambda_{n,d}^{-1},d))   \qquad \forall n,d \in \naturals.
\]
This implies that for the special choice 
\[
	n = \left\lceil 
	S_c \, h(cT(\varepsilon^{-1},d))  
	\right\rceil \ge 1
\]
	we have
	\[
	T(\lambda_{n,d}^{-s},d) \ge T(\varepsilon^{-1},d).
	\]
Since $T(\cdot,d)$ is non-decreasing, it follows that $\lambda_{n,d} \le \varepsilon$, and thus 
\[
\comp(\varepsilon,d) \le  \left\lceil 
S_c \, h(cT(\varepsilon^{-1},d))  
\right\rceil.  
\]

Consequently,
\begin{align}
\nonumber
\lim_{\varepsilon^{-1} + d \to \infty} \frac{\comp(\varepsilon,d)}{ h(2cT(\varepsilon^{-1},d))  }
& \le \lim_{\varepsilon^{-1} + d \to \infty} \frac{ \left\lceil 
	S_c \, h(cT(\varepsilon^{-1},d)) \right\rceil}{ [h(cT(\varepsilon^{-1},d))]^2  } \qquad \text{by \eqref{eq:h_cond}}
\\
\nonumber
& \le \lim_{\varepsilon^{-1} + d \to \infty} \frac{
	S_c \, h(cT(\varepsilon^{-1},d))+1}{ [h(cT(\varepsilon^{-1},d))]^2  }  \\
& = \lim_{\varepsilon^{-1} + d \to \infty} \left \{\frac{
	S_c }{ h(cT(\varepsilon^{-1},d))  }  + \frac{
	1}{ [h(cT(\varepsilon^{-1},d))]^2 } \right \} \label{eq:alastone}.
\end{align}	
The finiteness of $S_c$ implies that $\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d) = \infty$ for all $d \in \naturals$.  Together with the hypothesis on $\lim_{d \to \infty} T(\varepsilon^{-1},d)$, we can infer that $\lim_{\varepsilon^{-1} + d \to \infty} T(\varepsilon^{-1},d) = \infty$.  This implies that the limit in \eqref{eq:alastone} vanishes.  Since $c$ is arbitrary, we have satisfied the condition of Definition \ref{def:subhT}.

@@

\noindent \textbf{Necessary Condition:}\\
Suppose that the problem is  sub-$h \circ T$ tractable. Then, for any $c>0$ there exists a positive integer $V_c$ such that  
\[
	\comp(\varepsilon,d)\le 
	\left\lfloor h(cT(\varepsilon^{-1},d)) \right\rfloor \qquad \forall \varepsilon^{-1} + d \ge V_c.
\]
By the definition of complexity in \eqref{eq:comp_def}, it follows that 
\[
	\lambda_{n(\varepsilon,d,c),d} \le \varepsilon \quad \forall \varepsilon^{-1} + d \ge V_c, \qquad \text{where }
	n(\varepsilon,d,c):=\left\lfloor h(cT(\varepsilon^{-1},d)) \right\rfloor.
\]
Since $h(cT(\cdot,d))$ is non-decreasing, it follows that
\begin{equation}
		n(\varepsilon,d,c)\le  h(cT(\varepsilon^{-1},d)) \le  h(cT(\lambda^{-1}_{n(\varepsilon,d,c),d},d)) \qquad  \quad \forall \varepsilon^{-1} + d \ge V_c. \label{eq:nedcbound}
\end{equation}

For a fixed $c$ and $d$, define
\begin{equation} \label{eq:nstardef}
	\varepsilon^{-1}_{\max}(d,c) := \max(V_c - d,\lambda_{\max}^{-1}), \qquad n^*(d,c) : = n(\varepsilon^{-1}_{\max}(d,c),d,c) \le h(cT(\varepsilon^{-1}_{\max}(d,c),d)).
\end{equation} 
By varying $\varepsilon^{-1}$ in the interval $[\varepsilon^{-1}_{\max}(d,c),\infty)$,  the integer $n(\varepsilon,d,c)$ attains all values greater than or equal to $n^*(d,c)$.  Thus, by \eqref{eq:nedcbound},
\begin{align}
		n &\le h(cT(\lambda^{-1}_{n,d},d)) \qquad  \forall n \ge n^*(d,c), \label{eq:subhTneq}\\
		\nonumber
		\frac{1}{h(2 cT(\lambda^{-1}_{n,d},d)) } & \le \frac{1}{[h(cT(\lambda^{-1}_{n,d},d))]^{2}}  \le \frac 1{n^2}   \qquad  \forall n \ge n^*(d,c) \qquad \text{by \eqref{eq:h_cond}}. \\
		\intertext{Combining this inequality with the upper bound on $n^*(d,c)$ yields}
		\nonumber
		S_{2c} & =\sup_{d\in\naturals}  \, \sum_{i=1}^\infty \frac1{h(2cT(\lambda_{i,d}^{-1},d))} \\
		\nonumber
		&  = \sup_{d\in\naturals} \left [
		\sum_{i=1}^{n^*(d,c)-1} \frac1{h(2cT(\lambda_{i,d}^{-1},d))}
		+ \sum_{i=n^*(d,c)}^\infty \frac1{h(2cT(\lambda_{i,d}^{-1},d))} 
		\right]\\
		 \nonumber
		 & \le \sup_{d\in\naturals} \left [
\frac{n^*(d,c)-1}{h(2cT(\lambda_{\max}^{-1},d))}
+ \sum_{i=1}^\infty \frac1{n^2}
\right] 
 \qquad \text{by \eqref{eq:subhTneq}}\\ 
 \label{eq:lastone}
		 & \le \sup_{d\in\naturals} \left [
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}
		\right] + \frac{\pi^2}{6}
		\qquad \text{by \eqref{eq:nstardef}.}
\end{align}

Note that 
\[
\varepsilon^{-1}_{\max}(d,c) = \max(V_c - d,\lambda_{\max}^{-1}) 
= \begin{cases} V_c - d \le V_c, & d \in \{1, \ldots, d^*_c : = \lfloor V_c - \lambda_{\max}^{-1} \rfloor\} \\
\lambda_{\max}^{-1}, & d \in \{d^*_c+1, \ldots\}
\end{cases}
\]
Recalling that $T$ is non-decreasing in it's arguments, this implies that 
\begin{align*}
    \MoveEqLeft{\sup_{d\in\naturals} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}}\\
  &\le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*_c\}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}, 
  \sup_{d\in \{d^*_c+1, \ldots \}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \} \\
  & \le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*_c\}}  \,
		 \frac{ h(cT(V_c,d))}{h(2cT(\lambda_{\max}^{-1},d))}
  , 
  \sup_{d\in \{d^*_c+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \}
  \\
  & \le \max \left \{  
		 \frac{ h(cT(V_c,d^*_c))}{h(2cT(\lambda_{\max}^{-1},1))}, 
  \sup_{d\in \{d^*_c+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{[h(cT(\lambda_{\max}^{-1},d))]^2}
  \right \} \\
  & = \max \left \{  
		 \frac{ h(cT(V_c,d^*_c))}{h(2cT(\lambda_{\max}^{-1},1))}, 
		 \frac{1}{h(cT(\lambda_{\max}^{-1},d_c^*+1))}
  \right \} \\
  & < \infty.
\end{align*}
Combining this bound with \eqref{eq:lastone} establishes the finiteness of $S_{2c}$ and completes the proof of the necessary condition.

>>>>

Let $h:[0,\infty) \to [1,\infty)$ satisfy
\begin{equation} \label{eq:h_cond}
	h \text{ is strictly increasing}, \qquad 
	h(0)=1, \qquad  [h(x)]^c \le h(c x) \quad \forall x \ge 0, c > 0.
\end{equation}
\peter{Let us make the slightly stronger assumption on $h$ that
\[
h(x)\cdot h(y) \le h(x+y)\quad \forall x,y\in [0,\infty),
\]
and let us denote the inverse function of $h$ by $h^{-1}$. 
Then, for arbitrary $z,w \in [1,\infty)$ we choose $x=h^{-1}(z)$ and $y=h^{-1}(w)$, to obtain
\[
z w \le h (h^{-1}(z)+h^{-1}(w)),
\]
which yields by the monotonicity of $h^{-1}$ that
\begin{equation}\label{eq:submult_h_inverse}
h^{-1}(z w) \le h^{-1}(z)+h^{-1}(w).
\end{equation}
}

\fred{There is a problem here because I was assuming that 
\[
\lim_{x \to \infty} \frac{f(x)}{g(x)} = 0 \iff 
\lim_{x \to \infty} \frac{\log(f(x))}{\log(g(x))} = 0.
\]
But this is NOT the case, e.g., $f(x) = x^2$ and $g(x) = x^4$.  So we may need 
\begin{definition} \label{def:subhT}
	A problem is sub-$h \circ T$ tractable if
	\begin{equation*}
		\lim_{\varepsilon^{-1} + d \to \infty} \frac{h^{-1}(\comp(\varepsilon,d))}{T(\varepsilon^{-1},d,\vp)} = 0.
	\end{equation*}
\end{definition}
And NOT what I had below.}

\begin{definition} \label{def:subhT}
	A problem is sub-$h \circ T$ tractable if
	\begin{equation*}
		\lim_{\varepsilon^{-1} + d \to \infty} \frac{\comp(\varepsilon,d)}{h(T(\varepsilon^{-1},d,\vp))} = 0.
	\end{equation*}
\end{definition}

@@

\noindent \textbf{Sufficient Condition:}\\	
Suppose that $S_c <\infty$ for all $c>0$.  Since the terms of the series in the definition of $S_c$ are non-increasing, we have 
	\[
		\frac{n}{h(cT(\lambda_{n,d}^{-1},d))}\le S_c \qquad \forall n,d \in \naturals.
	\]
or equivalently
\[
	n\le S_c \, h(cT(\lambda_{n,d}^{-1},d))   \qquad \forall n,d \in \naturals.
\]
This implies that for the special choice 
\[
	n = \left\lceil 
	S_c \, h(cT(\varepsilon^{-1},d))  
	\right\rceil \ge 1
\]
	we have
	\[
	T(\lambda_{n,d}^{-s},d) \ge T(\varepsilon^{-1},d).
	\]
Since $T(\cdot,d)$ is non-decreasing, it follows that $\lambda_{n,d} \le \varepsilon$, and thus 
\[
\comp(\varepsilon,d) \le  \left\lceil 
S_c \, h(cT(\varepsilon^{-1},d))  
\right\rceil \qquad \forall c > 0.  
\]

\peter{
We could go on like this:

It is no loss of generality to assume that $S_c \ge 1$. From above we get 
\[
h^{-1}(\comp(\varepsilon,d)) \le h^{-1} \left( \left\lceil 
S_c \, h(cT(\varepsilon^{-1},d))  
\right\rceil \right) \le h^{-1} \left(  
2 S_c \, h(cT(\varepsilon^{-1},d))  
 \right) \qquad \forall c > 0,
\]
where we used that the argument of the ceiling function is at least one.

From this we deduce, by \eqref{eq:submult_h_inverse},
\[
h^{-1}(\comp(\varepsilon,d)) \le
 h^{-1} (2 S_c) + cT (\varepsilon^{-1},d).
\]
Now, for sufficiently large $\varepsilon^{-1}+d$, we must have
\[
c T (\varepsilon^{-1},d) \ge h^{-1} (2 S_c).
\]
This is because the finiteness of $S_c$ implies that $\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d) = \infty$ for all $d \in \naturals$,  together with the hypothesis on $\lim_{d \to \infty} T(\varepsilon^{-1},d)$.
So, 
\[
h^{-1}(\comp(\varepsilon,d)) \le 2 cT (\varepsilon^{-1},d).
\]
Since the latter holds for any $c>0$, we have sub-$h\circ T$-tractability.

}

Consequently, choosing $c=1/2$, it follows that 
\begin{align}
\nonumber
\lim_{\varepsilon^{-1} + d \to \infty} \frac{\comp(\varepsilon,d)}{ h(T(\varepsilon^{-1},d))  }
& = \lim_{\varepsilon^{-1} + d \to \infty} \frac{\comp(\varepsilon,d)}{ h(2cT(\varepsilon^{-1},d))  } \\
& \le \lim_{\varepsilon^{-1} + d \to \infty} \frac{ \left\lceil 
	S_c \, h(cT(\varepsilon^{-1},d)) \right\rceil}{ [h(cT(\varepsilon^{-1},d))]^2  } \qquad \text{by \eqref{eq:h_cond}}
\\
\nonumber
& \le \lim_{\varepsilon^{-1} + d \to \infty} \frac{
	S_c \, h(cT(\varepsilon^{-1},d))+1}{ [h(cT(\varepsilon^{-1},d))]^2  }  \\
& = \lim_{\varepsilon^{-1} + d \to \infty} \left \{\frac{
	S_c }{ h(cT(\varepsilon^{-1},d))  }  + \frac{
	1}{ [h(cT(\varepsilon^{-1},d))]^2 } \right \} \label{eq:alastone}.
\end{align}	
The finiteness of $S_c$ implies that $\lim_{\varepsilon \to 0} T(\varepsilon^{-1},d) = \infty$ for all $d \in \naturals$.  Together with the hypothesis on $\lim_{d \to \infty} T(\varepsilon^{-1},d)$, we can infer that $\lim_{\varepsilon^{-1} + d \to \infty} T(\varepsilon^{-1},d) = \infty$.  This implies that the limit in \eqref{eq:alastone} vanishes.

@@

\noindent \textbf{Necessary Condition:}\\
Suppose that the problem is  sub-$h \circ T$ tractable. Then, for any $c>0$ there exists a positive integer $V_c$ such that  \fred{having a little trouble removing the $c$ here.} \peter{Here we do not need to remove the $c$, and actually shouldn't. Just in the limit relation it was not necessary.}
\[
	\comp(\varepsilon,d)\le 
	\left\lfloor h(cT(\varepsilon^{-1},d)) \right\rfloor \qquad \forall \varepsilon^{-1} + d \ge V_c.
\]
By the definition of the information complexity in \eqref{eq:comp_def}, it follows that 
\[
	\lambda_{n(\varepsilon,d,c),d} \le \varepsilon \quad \forall \varepsilon^{-1} + d \ge V_c, \qquad \text{where }
	n(\varepsilon,d,c):=\left\lfloor h(cT(\varepsilon^{-1},d)) \right\rfloor \peter{+1}.
\]
Since $h(cT(\cdot,d))$ is non-decreasing, it follows that
\begin{equation}
		n(\varepsilon,d,c)\le  h(cT(\varepsilon^{-1},d)) \peter{+1}\le  h(cT(\lambda^{-1}_{n(\varepsilon,d,c),d},d)) \peter{+1}\qquad  \quad \forall \varepsilon^{-1} + d \ge V_c. \label{eq:nedcbound}
\end{equation}

For a fixed $c$ and $d$, define
\begin{equation} \label{eq:nstardef}
	\varepsilon^{-1}_{\max}(d,c) := \max(V_c - d,\lambda_{\max}^{-1}), \qquad n^*(d,c) : = n(\varepsilon^{-1}_{\max}(d,c),d,c) \le h(cT(\varepsilon^{-1}_{\max}(d,c),d)) \peter{+1}.
\end{equation} 
By varying $\varepsilon^{-1}$ in the interval $[\varepsilon^{-1}_{\max}(d,c),\infty)$,  the integer $n(\varepsilon,d,c)$ attains all values greater than or equal to $n^*(d,c)$.  Thus, by \eqref{eq:nedcbound},
\begin{align}
		n &\le h(cT(\lambda^{-1}_{n,d},d)) \qquad  \forall n \ge n^*(d,c), \label{eq:subhTneq}\\
		\nonumber
		\frac{1}{h(2 cT(\lambda^{-1}_{n,d},d)) } & \le \frac{1}{[h(cT(\lambda^{-1}_{n,d},d))]^{2}}  \le \frac 1{n^2}   \qquad  \forall n \ge n^*(d,c) \qquad \text{by \eqref{eq:h_cond}}. \\
		\intertext{Combining this inequality with the upper bound on $n^*(d,c)$ yields}
		\nonumber
		S_{2c} & =\sup_{d\in\naturals}  \, \sum_{i=1}^\infty \frac1{h(2cT(\lambda_{i,d}^{-1},d))} \\
		\nonumber
		&  = \sup_{d\in\naturals} \left [
		\sum_{i=1}^{n^*(d,c)-1} \frac1{h(2cT(\lambda_{i,d}^{-1},d))}
		+ \sum_{i=n^*(d,c)}^\infty \frac1{h(2cT(\lambda_{i,d}^{-1},d))} 
		\right]\\
		 \nonumber
		 & \le \sup_{d\in\naturals} \left [
\frac{n^*(d,c)-1}{h(2cT(\lambda_{\max}^{-1},d))}
+ \sum_{i=1}^\infty \frac1{n^2}
\right] 
 \qquad \text{by \eqref{eq:subhTneq}}\\ 
 \label{eq:lastone}
		 & \le \sup_{d\in\naturals} \left [
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}
		\right] + \frac{\pi^2}{6}
		\qquad \text{by \eqref{eq:nstardef}.}
\end{align}

Note that 
\[
\varepsilon^{-1}_{\max}(d,c) = \max(V_c - d,\lambda_{\max}^{-1}) 
= \begin{cases} V_c - d \le V_c, & d \in \{1, \ldots, d^*_c : = \lfloor V_c - \lambda_{\max}^{-1} \rfloor\} \\
\lambda_{\max}^{-1}, & d \in \{d^*_c+1, \ldots\}
\end{cases}
\]
Recalling that $T$ is non-decreasing in its arguments, this implies that 
\begin{align*}
    \MoveEqLeft{\sup_{d\in\naturals} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}}\\
  &\le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*_c\}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}, 
  \sup_{d\in \{d^*_c+1, \ldots \}} \,
		 \frac{ h(cT(\varepsilon^{-1}_{\max}(d,c),d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \} \\
  & \le \max \left \{  
  \sup_{d \in \{1, \ldots, d^*_c\}}  \,
		 \frac{ h(cT(V_c,d))}{h(2cT(\lambda_{\max}^{-1},d))}
  , 
  \sup_{d\in \{d^*_c+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{h(2cT(\lambda_{\max}^{-1},d))}
  \right \}
  \\
  & \le \max \left \{  
		 \frac{ h(cT(V_c,d^*_c))}{h(2cT(\lambda_{\max}^{-1},1))}, 
  \sup_{d\in \{d^*_c+1, \ldots \}}  \,
		 \frac{ h(cT(\lambda_{\max}^{-1},d))}{[h(cT(\lambda_{\max}^{-1},d))]^2}
  \right \} \\
  & = \max \left \{  
		 \frac{ h(cT(V_c,d^*_c))}{h(2cT(\lambda_{\max}^{-1},1))}, 
		 \frac{1}{h(cT(\lambda_{\max}^{-1},d_c^*+1))}
  \right \} \\
  & < \infty.
\end{align*}
Combining this bound with \eqref{eq:lastone} establishes the finiteness of $S_{2c}$ and completes the proof of the necessary condition.
